The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2018-06-04 Paper/jaccardbaetica.tex
review

The majority of olive oil amphorae produced in this province and shipped thought the provinces belong to Dressel 20 type \citep{dressel,martin-kilcher_romischen_1994}. This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}. Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in graffities with different information. However, graffities and \textit{tituli picti} have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}. By contrast, stamps are the most studied in this type of amphorae. Frequently, they were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}. 
The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria. Stamps was mostly formed for a code of three letters. These letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.

INCLUIR AQUÍ QUE LAS STAMPS VAN CAMBIANDO DE FECHA Y SIGLOS

@@

From the database, we collected the site where it was found and the stamp code. We also created a new column with the area where stamps were found (decir berni). Those areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire. Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalis} (currently Seville), \textit{Corduba} (currently C\'ordoba) and \textit{Astigi} (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .

@@

The results of the case study could be interpreted due to several reason according to previous hypotheses. On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR). This hypotheses would explain the fact that we do not find similar stamps in different workshops, however, we found different amphora stamps in a same workshops that they would be barely difficult to assign different owners. Neither a kind of quality distinction does seem to be used to specify the quality or value of the product \citep{callender}. On the other hand, it can be interpreted somehow a batch systematic organization (CITAR). Considering that Dressel 20 was not marked in several cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped (CITAR JUAN MOROS). This method would be used as an identifier to count the number of amphorae of a branch (CITAR). This it could be explain wherefore we found different stamps in a same workshop. 
In any case, we do not have enough archaeological evidence that can validated the interpretations presented here and our results are certainly valid only with the context of our case study.

>>>>

The majority of olive oil amphorae produced in this province and shipped thought the provinces belong to Dressel 20 type \citep{dressel,martin-kilcher_romischen_1994}. This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}. Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in graffities with different information. However, graffities and \textit{tituli picti} have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}. By contrast, stamps are the most studied in this type of amphorae. During almost three centuries, stamps were used to mark amphorae with a different chronological frequency \citep{remesal_sellar_2016}. Frequently, they were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}. 
The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria. Stamps was mostly formed for a code of three letters. These letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.

@@

From the database, we collected the site where it was found and the stamp code. We also created a new column with the area where stamps were found. These areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire. Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalis} (currently Seville), \textit{Corduba} (currently C\'ordoba) and \textit{Astigi} (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .

@@

The results of the case study could be interpreted due to several reason according to previous hypotheses. On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR). This hypotheses would explain the fact that we do not find similar stamps in different workshops, however, we found different amphora stamps in a same workshops that they would be barely difficult to assign different owners. Neither a kind of quality distinction does seem to be used to specify the quality or value of the product \citep{callender}. On the other hand, it can be interpreted somehow a batch systematic organization. Considering that Dressel 20 was not marked in several cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}. This method would be used as an identifier to count the number of amphorae of a branch (CITAR). This it could be explain wherefore we found different stamps in a same workshop. 
In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are certainly valid only with the context of our case study.
